---
title:
date: "January 15, 2019"
output:
  rmdformats::html_clean:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
               cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Syllabus (FIN 685) Seminar in Finance
<br>
<br>
<center>
<b>SUNY POLYTECHNIC INSTITUTE</b>

<b>SCHOOL OF BUSINESS ADMINISTRATION</b>
</center>
<br>
<br>
**Instructor**: Matthew Brigida, Ph.D.

**Office**: Donovan 1277

<!-- **Office Hours**: Tuesday 4:00--6:00 and Monday/Wednesday 11:00--12:00 -->

**Email**:  matthew.brigida@sunypoly.edu

**Class Location**:  Online (Blackboard)

<!-- **Text**: [*Investments* by Bodie, Kane, and Marcus, 8th edition (or later), ISBN: 0077261453](http://www.amazon.com/Investments-8th-Edition-Zvi-Bodie/dp/0077261453/ref=sr_1_5?ie=UTF8&qid=1407685749&sr=8-5&keywords=investment+bodie+kane+marcus) -->

**Supplementary Material**: [5MinuteFinance.org](https://www.5minutefinance.org/)

## Course Structure and Interaction

-  All interaction will be through Blackboard, with a majority taking place on the discussion boards and via BlackBoard Collaborate vidoes.  Since we may not be able to meet at the same time, you'll want to check the `Recordings` section of Collaborate.

-  The Blackboard discussion boards should generally be used for any course questions.  Only use email for questions that contain personal information.

-  There may be some useful videos at the following links: [Investments](https://www.youtube.com/playlist?list=PLRK1VeglPb4NO-D0HOn0QE9lQlBNR1vHK) and [Portfolio Theory](https://www.youtube.com/playlist?list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6).  I'll let you know if I post anything particularly useful here.


## Academic Honesty Policy

Academic dishonesty will not be tolerated in this class. Cheating
on quizzes, examinations, and other forms of dishonesty (e.g., plagiarism, collusion, and
falsification of data) will be dealt with in a serious and formal manner. The penalty for academic
dishonesty in this class will be course failure. That is, any student who is found to be cheating
or engaged in other academically dishonest behavior will be failed for this course for this
semester. Course withdrawals to avoid such a failure will not be permitted. As a student, you
have a responsibility to become familiar with the Academic Honesty Policy found in the Student
Rights, Regulations, and Procedures Handbook.

## Online Office Hours

I will be logged on the course throughout the week to answer questions and review the discussions.  If a few students have questions on the material, and there is a good time to meet, I'll log into BB Collaborate and record the answers.

Students do not need to access the course simultaneously, and therefore have a measure of flexibility regarding when they access the course.  However, students should access the course regularly to keep up with assignments, and to maintain a presence in the Discussion Room.  Students should also often check their SUNY Poly email account.

## Description

From the [Course Catalog](https://webapp.sunypoly.edu/grad-catalog-2017-2018/courses/finance/):

> An integrating experience to apply the varied skills and knowledge accumulated through the required course work to make the student competitive in capital markets.  Special emphasis will be upon mastery of body of accounting and financial knowledge including significant current development on the economic and financial scene.  Students acquire greater understanding of global capital markets, demonstrate the ability to use the tools and techniques of accounting and investment analysis in the valuation of assets, and provide a synthesis of all previous related course work.  Prerequisites:  ACC 520 and FIN 525.

## Topics Covered

<!-- some good papers here: https://www2.bc.edu/thomas-chemmanur/page/mf891_syllabus99.html -->

-  Capital Structure
    + In Perfect Capital Markets
        - [Modigliani, F. and M. Miller "The Cost of Capital, Corporation Finance and the Theory of Investment" American Economic Review, June 1958, 261-297.](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C33&q=Modigliani%2C+F.+and+M.+Miller+%E2%80%9CThe+Cost+of+Capital%2C+Corporation+Finance+and+the+Theory+of+Investment%E2%80%9D+American+Economic+Review%2C+June+1958%2C+261-297.&btnG=)
    + With Taxes
	    - [Miller, M., "Debt and Taxes," Journal of Finance, June 1977, 32, 261-276.](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1540-6261.1977.tb03267.x) 
	+ With Agency Costs
	    - [Jensen, M. and W. Meckling, "Theory of the Firm: Managerial Behavior, Agency Costs, and Ownership Structure," Journal of Financial Economics, October 1976, 3, 305-360.](https://www2.bc.edu/thomas-chemmanur/phdfincorp/MF891%20papers/Jensen%20and%20Meckling%201976.pdf)
- Adverse Selection
   + [Ackerlof, G. A., "The market for lemons: Quality Uncertainty and the Market Mechanism," The Rand Journal of Economics.](https://pages.ucsd.edu/~aronatas/project/academic/Akerlof%20on%20Lemons.pdf)
- Adverse Selection and Capital Structure
   + [Myers, S. and N. Majluf, "Corporate Financing and Investment Decisions When Firms Have Information that Investors Do Not Have," Journal of Financial Economics, June 1984,187-221.](https://www.nber.org/papers/w1396.pdf)
-  Dividend Policy
    + [Black, F., "The Dividend Puzzle", The Journal of Portfolio Management, (Dec 1996): 8-12.](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C33&q=fischer+black+dividend+puzzle&btnG=)
    <!-- + Miller, M. and K. Rock, "Dividend Policy Under Asymmetric Information," Journal of Finance, September 1985, 40, 1031-1051.  -->
-  The Market for Corporate Control
    + [Size, ownership and the market for corporate control---Robinson, David T., J. Corporate Finance 2009]()
    + IPOs/SEOs
      - [A review of IPO activity, pricing, and allocations](https://www.nber.org/papers/w8805.pdf)
      - [Why has IPO underpricing changes over time?](https://scholar.google.com/scholar?cluster=2692753777902135925&hl=en&as_sdt=0,33)
      <!---  - [Rock, K., 1986, Why new issues are underpriced, Journal of Financial Economics 15, 187-212.]() -->
    + Mergers and Acquisitions
- Security Design
    + [Harris, Milton and Artur Raviv (1989), "The Design of Securities," Journal of Financial Economics, 24, 255-287.]()
-  Market Structure
    + [The Only Game in Town](Bagehot)
    + [High frequency market microstructure](http://dx.doi.org/10.1016/j.jfineco.2015.01.003)
-  Banking post 2008
    + [Bank capital and lending relationships](https://scholar.google.com/scholar?cluster=2931503161409199966&hl=en&as_sdt=0,33)
<!-- -  Contingent Claim Analysis in Complete Markets -->
<!--     + []() -->
-  Artificial Intelligence and Machine Learning in Accounting and Finance (See the McKiney documents in BB)
<!-- use mckinsey docs -->
    + Uses by big 4 accounting firms
    + Loan Origination
    + Mergers and Acquisitions/Corporate Strategy
    + Valuation
	
	
## Research Paper Discussion

How to summarize a paper:

#. What is the purpose of the paper?  What is the unanswered question in the literature?
#. What analysis method did the author use? Did they use an empirical (statistics), theoretical, or survey approach?
#. What was their conclusion?
#. What are the limitations to their conclusion? What sort of assumption did they have to make? 
#. Can you think of further research based on the paper? 

##  Possible Paper Topic Areas

Below are a set of possible paper topics.  You do not have to choose one of these topics---in fact many of these are simply topics I am interested in but can't pursue for various reasons.  However, reading through these topics should at the very least give you a good idea of the types and breadth of topics you can pursue.

-  [Capital Structure and Lending Policy at US Banks---FDIC Data Set](#banking-analyses)
-  [Electricity Markets---Real Option Valuation of Power Plants, Renewable Power Integration](#electricity-markets)
-  [High-Frequency Trading](#high-frequency-trading)
-  [Energy Markets](#energy-markets)
-  [Crypotocurrency Markets](#cryptocurrency-markets)
-  [A case study outlining the effect of AI/ML in Accounting and Finance](#ai/ml-in-finance-and-accounting)

## Course Communication

All communication will be through Blackboard and email. If you have question you are strongly encouraged to post it to a discussion forum instead of emailing me.  This way, other students can get the benefit of the question and answer.  It also saves me from answering the same question many times via email, and frees me up to answer more questions and generally provide more effective instruction for you.

## Grading

<center>

```{r grading, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Item | Points |
| ---- | ------ |
| Paper Summaries | 60 |
| Research Paper |  40 |
| Total Points | 100 |
"
cat(tabl)
```

</center>

<center>
**Final grades will be assigned according to the following scale**:
</center>

- 90 - 100 A
- 80 - 89.9 B
- 70 - 79.9 C
- 60 - 69.9 D
- $<$ 60 F

> +/- grades may be assigned at the instructors discretion.

## Tentative Outline/Schedule

- 8/27: Course Intro
- 9/3:
- 9/10: Capital Structure Summaries
- 9/17: 
- 9/24: Adverse Selection and Dividend Policy Summaries
- 10/1: 
- 10/8: The Market for Corporate Control Summaries 
- 10/15: 
- 10/22: Security Design and Market Microstructure and Banking
- 10/29: 
- 11/5: AI
- 11/12: 
- 11/19: 
- 11/26--12/8: **The last day to submit your papers.**
- 12/10: Finals Week (nothing due)






<!-- Spring -->
<!-- - 1/15: Chapter 1 -->
<!-- - 1/22:  Chapter 2 \& 3 -->
<!-- - 1/29: Chapter 3 \& 4 (Note: chapter 3 is more important than 4.) **Start Market Depth Project** -->
<!-- - 2/5: Chapter 5 Some useful videos for chapters 5, 6, and 7 are [here](https://www.youtube.com/watch?v=iyxoZZUULCQ&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=4), [here](https://www.youtube.com/watch?v=ZtxvkdThpck&index=3&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6), [here](https://www.youtube.com/watch?v=XX1h196RLUg&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=2), and [here](https://www.youtube.com/watch?v=dpB5fx5lbA8&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=1). -->
<!-- - 2/12: Chapter 6 -->
<!-- - 2/19: Chapter 7 -->
<!--   + **Recommended <font color="red">Due Date</font> for the Market Depth Project** -->
<!--   + **Start MV Optimal Portfolio Project** [See This Video](https://www.youtube.com/watch?v=HQrba-gn1Gs&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=10) and [These Spreadsheet Instructions](https://www.youtube.com/watch?v=jjjQV__bSHk&t=94s&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=12) -->
<!-- - 2/26: [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/) -->
<!-- - *Midterm is through Securitization.  Final exam will cover from here onward.* -->
<!-- - 3/12: Interlude: The term structure of interest rates and measuring bond price sensitivity to interest rates. **No calculations required** -->
<!-- - 3/19: Chapter 20 -->
<!-- - 3/26: Chapter 21 -->
<!--   + **Recommended <font color="red">Due Date</font> for the MV Optimal Portfolio Project** -->
<!--   + **Start Option Pricing (Black-Scholes) Project** -->
<!-- - 4/1: Chapter 22 -->
<!-- - 4/8: Chapter 23 -->
<!-- - 4/15: Projects/Trade -->
<!-- - 4/22: Projects/Trade -->
<!--   + **<font color="red">All assignments/exams/projects due by the end of the week</font>.** -->
<!-- - 5/6: Finals Week -->

<!-- Summer -->
<!-- - 5/20: Chapters 1 and 2 -->
<!-- - 5/27: Chapter 3 \& 4 (Note: chapter 3 is more important than 4.) **Start Market Depth Project** -->
<!-- - 6/3: Chapter 5 Some useful videos for chapters 5, 6, and 7 are [here](https://www.youtube.com/watch?v=iyxoZZUULCQ&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=4), [here](https://www.youtube.com/watch?v=ZtxvkdThpck&index=3&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6), [here](https://www.youtube.com/watch?v=XX1h196RLUg&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=2), and [here](https://www.youtube.com/watch?v=dpB5fx5lbA8&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=1). -->
<!-- - 6/10: Chapters 6 and 7 -->
<!--   + **Recommended <font color="red">Due Date</font> for the Market Depth Project** -->
<!--   + **Start MV Optimal Portfolio Project** [See This Video](https://www.youtube.com/watch?v=HQrba-gn1Gs&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=10) and [These Spreadsheet Instructions](https://www.youtube.com/watch?v=jjjQV__bSHk&t=94s&list=PLRK1VeglPb4MeP83AIOZFj2quPZ1t4dN6&index=12) -->
<!-- - 6/17: [Securitization](https://matt-brigida.github.io/securitization-presentation-abridged/) -->
<!--   + *Midterm is through Securitization.  Final exam will cover from here onward.* -->
<!-- - 6/24: Chapters 20 and 21 -->
<!--   + **Recommended <font color="red">Due Date</font> for the MV Optimal Portfolio Project** -->
<!--   + **Start Option Pricing (Black-Scholes) Project** -->
<!-- - 7/1: Chapters 22 and 23 -->
<!-- - 7/8: Projects/Trade -->
<!--   + **<font color="red">All assignments/exams/projects due by the end of the week</font>.** -->


<!-- > Based on student input, we may add topics in portfolio theory such as those found in chapters 5 through 10, and chapter 24. -->

## General Notes

- All times referred to in this course are Eastern Standard---unless otherwise indicated.
- There will be no make up or extra points assignments.
- Plaigarism will result in prosecution to the fullest extent possible under university rules.
- You are responsible for material covered in the online discussion, as well as text material.
-  **Internet Access:** This course requires that you have regular access to the internet to submit work.  You should not take this course if you plan on being in an area with insufficient internet access. "My internet was down for a week" is not an acceptable reason to hand in late work.
-  **Adding or Dropping the Course:** To add or drop the course the student should consult the university guidelines and withdrawal dates (as well as thier advisor). The course instructor is not involved in a student's adding or withdrawing from the course.
- **Software:** You may use word processing and spreadsheet software in
this course. Common examples of such software are Microsoft Word and
Excel. However, there is no need to buy this software if you don't already
have it. There are many free (open-source) alternatives which are just as
good (and which allow you to save/read files as .doc(x), .pdf, and .xls(x)).
For example I use LibreOffice (http://www.libreoffice.org). 



## Banking Analyses

I have created this [panel data set]() which I [used in this paper]().  You can use this data set to estimate the relationship between a bank's capital structure and aspects of its loan portfolio or performance.

-  Do higher levels of equity in a banks capital structure lead to lower profitability?  This analysis would be best done in R or Python instead of Excel.

You may want to look at the paper, "Bank Capital and Lending Relationships"  (and the papers that cited it) prior to starting your analysis.

## High Frequency Trading (HFT)

- Survey the effect of Regulation NMS on market structure and the rise of HFT.

- Statistical analyses to identify, for example, the impact of HFT on market efficiency.

## Cryptocurrency Markets

There are a couple very interesting aspects to cryptocurrencies and cryptocurrency markets.

-  Relationships between the behavior of various currencies.  This is an exercise in time series analysis, and to be done correctly you should use the idea of cointegration.

-  An analysis of the extent of asymmetric information in cryptocurrency markets and the effect on market viability.

-  An analysis of [Libra]()---how does it differ from other currencies and how is it similar? 

-  A survey of the various trading venues for cryptocurrencies and their derivatives.

-  A survey of various cryptocurrency market APIs.

## Electricity Markets

- A survey of the various electricity market structures around the US.

- How can we incentivize overcapacity within a deregulated utility structure?  Survey the problem and possible solutions, hopefully adding ideas of your own.

-  Survey the percent renewables by ISO and the possible effect on market prices.

## Energy Markets

-  Survey the literature on the relationship between natural gas and crude oil prices.

-  Survey accounting of oil reserves (explaining various definitions of 'proven').

## AI/ML in Finance and Accounting

The industry with the largest investment in Artificial Intelligence and Machine Learning (AI/ML) is Financial and Accounting Services (cite McKinsey document).


## BRTI Tick Data {#brti}

Below are millisecond data for the Bitcoin Real-Time Index on Jan 3, 2018.  The data are pulled via the IB API.  With data of this type, you can dicover interesting features of market data---for example, notice the "last trade" is often outside of the bid/offer.  This is particularly important to understand when trading illiquid contracts, such as some options.  You can left-click your mouse, and drag, to zoom in on subintervals.

Using such data is not necesary in this course, though if you would like to use similar data in your analysis let me know.  I can either pull the data you want for you, or I can request an account for you where you can pull the data.

```{r error=FALSE, warning=FALSE, eval = TRUE}
library(dygraphs)

book <- readRDS("./jan_fut_12_18_book.rds")

dygraph(book['2018-01-03 09:30:00/2018-01-03 16:00:00'])

```

## Some Notes on Spreadsheet Design

The following are notes from another course I teach.  You likely will not have to submit any spreadsheets in this course.  However, I'll leave these notes here because there is certainly no harm in reading (or rereading) them. 

You should construct your spreadsheet as if you were an analyst at a company, and you were going to submit the spreadsheet to upper management.  Therefore, getting the correct answer can be considered the minimal amount of work.  The spreadsheet should be easily readable and organized.  There are a couple of reasons why this is important: (1) management often will check some numbers (or maybe change a few inputs if they have more up to date information) and it will reflect very poorly on you if they have to search around through a muddled and ill-conceived spreadsheet; and (2) anyone should be able to pick up your spreadsheet and complete it if you are not there (vacation, sick, or hopefully promoted).  Following are a couple tips on spreadsheet design, though it is far from exhaustive.

- Hard-code as little as possible.  You want a few cells for your inputs, or a place where you put your data, and then every other cell is linked and feeds off of these input cells. This way, to update your spreadsheet you simply change the inputs or drop in new data.
- Take the time to label cells, and put in appropriate comments if necessary  - though comments should not be used excessively. Also, it is common to change the cell color depending on whether it is hard-coded (an input) or a formula.  This way you (or anyone else) can immediately look at a cell and tell whether it is one in which you can type (an input).  Don't forget to include a key.
- It is often better to add tabs to a spreadsheet than continue calculations on one tab.  You can easily page through spreadsheet tabs with `Ctrl+Shift` and `Page-up` or `Page-down`.
- Pivot tables.  While we probably won't need them in this course, you should nonetheless get to know them.  Pivot tables are incredibly useful for summarizing data, and it is very possible you will be asked in an interview whether you are familiar with them.  Similarly, get to know VLOOKUP.
-  If you are inputting a long formula, then break the calculation into multiple cells.  This makes it much easier to tell where a mistake was made - and everyone always spends a fair amount of time looking for errors.
- Excel has many built in formulas which can be useful, however it is important that you understand what the formula is doing to use them.  Blindly applying a formula can lead to trouble.  For example, if you use the IRR() function on cash flows with multiple roots, the formula will return the first root it finds without signaling to you that there are other roots.  Also, there are Excel formulas that are flat out incorrect - in particular the NPV() function.  So, use a function if it saves time, but first be sure you know what the function is doing and verify it works.  That said, in my experience it is better (and faster) to input your own formula instead of using Excel's.  You often have to break the calculation into a couple of steps, but this can be done quickly, and the result is a spreadsheet that you know works and is easily auditable.
